---
title: 'Data Analysis Assignment #1 (50 points total)'
author: "Hayashi, Dylan"
output:
  html_document: default
---

```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```


Submit both the .Rmd and .html files for grading. You may remove the instructions and example problem above, but do not remove the YAML metadata block or the first, "setup" code chunk.  Address the steps that appear below and answer all the questions. Be sure to address each question with code and comments as needed.  You may use either base R functions or ggplot2 for the visualizations.

-----

The following code chunk will:

(a) load the "ggplot2", "gridExtra" and "knitr" packages, assuming each has been installed on your machine,
(b) read-in the abalones dataset, defining a new data frame, "mydata,"
(c) return the structure of that data frame, and
(d) calculate new variables, VOLUME and RATIO.

Do not include package installation code in this document. Packages should be installed via the Console or 'Packages' tab. You will also need to download the abalones.csv from the course site to a known location on your machine. Unless a *file.path()* is specified, R will look to directory where this .Rmd is stored when knitting.

```{r analysis_setup1, message = FALSE, warning = FALSE}

# a) Load the ggplot2 and gridExtra packages.

library(ggplot2)
library(gridExtra)
library(knitr)

# b) Use read.csv() to read the abalones.csv into R, assigning the data frame to "mydata."

mydata <- read.csv("abalones.csv", sep = ",", stringsAsFactors = TRUE)

# c) Use the str() function to verify the structure of "mydata." You should have 1036 observations
# of eight variables.

str(mydata)

# d) Define two new variables, VOLUME and RATIO. Use the following statements to define VOLUME and
# RATIO as variables appended to the data frame "mydata."

mydata$VOLUME <- mydata$LENGTH * mydata$DIAM * mydata$HEIGHT
mydata$RATIO <- mydata$SHUCK / mydata$VOLUME

```

-----

### Test Items starts from here - There are 6 sections - Total 50 points

***##### Section 1: (6 points) Summarizing the data.***

(1)(a) (1 point) Use *summary()* to obtain and present descriptive statistics from mydata.  Use *table()* to present a frequency table using CLASS and RINGS. There should be 115 cells in the table you present.  

```{r Part_1a}
summary(mydata)
CLASS <- mydata$CLASS
RINGS <- mydata$RINGS
table(CLASS,RINGS)

```

**Question (1 point):  Briefly discuss the variable types and distributional implications such as potential skewness and outliers.**

***Answer: This dataset contains 8 native variables. Two of these variables, sex and class, are qualitative. The rest of the variables are quantitative. Additionally, there two derived variables, volume and ratio, that have been added. Sex appears to be roughly equally distributed among its three values. Class appears to have a bell (but not necessarily normal) shaped distribution. Analysis of histograms for each quantitative variable reveal that all variables aside from Height (which appears to resemble normal) have skewness. Length and Diameter are skewed left, while Whole Weight, Shuck, Rings, Volume, and Ratio are skewed right. Analysis of boxplots for each quantiative variable reveal that all variables have outliers. Rings, Ratio, Shuck, and Whole Weight have outliers above the mean, with the listing of variables presenting approximate descending frequency. Length, Diameter, Height, Rings, and Ratio also have outliers below the mean, with the listing of variables also presenting approximate descending frequency***

(1)(b) (1 point) Generate a table of counts using SEX and CLASS. Add margins to this table (Hint: There should be 15 cells in this table plus the marginal totals. Apply *table()* first, then pass the table object to *addmargins()* (Kabacoff Section 7.2 pages 144-147)).  Lastly, present a barplot of these data; ignoring the marginal totals.  

```{r Part_1b}
addmargins(table(mydata$SEX,mydata$CLASS))
barplot(table(mydata$SEX,mydata$CLASS),
       main = 'Class Frequency, Differentiated by Sex',
       xlab = 'Class',
       ylab = 'Frequency',
       col = c('Green','Blue','Purple'),
       legend = TRUE,
       beside = TRUE
       )
```

**Essay Question (2 points):  Discuss the sex distribution of abalones.  What stands out about the distribution of abalones by CLASS?**

***Answer: As expected, the frequency of infant abalones decreases as rings increases. Throughout all classes, there appears to be a greater number of male adult abaloned than female adult aboloned, with the difference becoming less pronounced as class (age) increase. Because the sex of infant abalones are not identified, it is impossible to tell the frequency ratio of males to females in infant abalone.***

(1)(c) (1 point) Select a simple random sample of 200 observations from "mydata" and identify this sample as "work." Use *set.seed(123)* prior to drawing this sample. Do not change the number 123. Note that *sample()* "takes a sample of the specified size from the elements of x." We cannot sample directly from "mydata." Instead, we need to sample from the integers, 1 to 1036, representing the rows of "mydata." Then, select those rows from the data frame (Kabacoff Section 4.10.5 page 87). 

Using "work", construct a scatterplot matrix of variables 2-6 with *plot(work[, 2:6])* (these are the continuous variables excluding VOLUME and RATIO). The sample "work" will not be used in the remainder of the assignment. 

```{r Part_1c}
set.seed(123)
work <- mydata[sample(1:nrow(mydata),20, replace = FALSE),]
plot(work[,2:6])
```

-----

***##### Section 2: (5 points) Summarizing the data using graphics.***

(2)(a) (1 point) Use "mydata" to plot WHOLE versus VOLUME. Color code data points by CLASS.

```{r Part_2a}
plot(mydata$VOLUME,mydata$WHOLE,
     main = 'Whole vs. Volume, Differentiated by Class',
     xlab = 'Whole',
     ylab = 'Volume',
     col = c('Gray','Blue','Purple','Green','Orange')[mydata$CLASS])

legend(x = 900,
       y = 200,
       legend = levels(mydata$CLASS),
       col = c('Gray','Blue','Purple','Green','Orange')[levels(mydata$CLASS)],
       pch = 1)
```


(2)(b) (2 points) Use "mydata" to plot SHUCK versus WHOLE with WHOLE on the horizontal axis. Color code data points by CLASS. As an aid to interpretation, determine the maximum value of the ratio of SHUCK to WHOLE.  Add to the chart a straight line with zero intercept using this maximum value as the slope of the line. If you are using the 'base R' *plot()* function, you may use *abline()* to add this line to the plot. Use *help(abline)* in R to determine the coding for the slope and intercept arguments in the functions. If you are using ggplot2 for visualizations, *geom_abline()* should be used.

```{r Part_2b}
##### FINISH #####
plot(mydata$WHOLE,mydata$SHUCK,
     main = 'Whole vs. Shuck, Differentiated by Class',
     xlab = 'Whole',
     ylab = 'Shuck',
     col = c('Gray','Blue','Purple','Green','Orange')[mydata$CLASS],
     abline(0,max(mydata$SHUCK / mydata$WHOLE)))

legend(x = 275,
       y = 100,
       legend = levels(mydata$CLASS),
       col = c('Gray','Blue','Purple','Green','Orange'),
       pch = 1)
```

**Essay Question (2 points):  How does the variability in this plot differ from the plot in (a)?  Compare the two displays.  Keep in mind that SHUCK is a part of WHOLE.  Consider the location of the different age classes.**

***Answer: (These two plots indicate that whole weight is positively correlated with both volume and shuck. The plot of Whole vs. Volume has greater variability than the plot of Whole vs. Shuck. Additionally, both plots show that the variability of each ratio (whole : volume and whole : shuck) tend to increase with age.)***

-----

***### Section 3: (8 points) Getting insights about the data using graphs.***

(3)(a) (2 points) Use "mydata" to create a multi-figured plot with histograms, boxplots and Q-Q plots of RATIO differentiated by sex. This can be done using *par(mfrow = c(3,3))* and base R or *grid.arrange()* and ggplot2. The first row would show the histograms, the second row the boxplots and the third row the Q-Q plots. Be sure these displays are legible.  

```{r Part_3a}
par(mfrow = c(3,3))
# Histograms
hist(mydata$RATIO[mydata$SEX=="F"],
     main = 'Histogram of Ratio, Sex = F',
     xlab = 'Ratio',
     col = 'Green')
hist(mydata$RATIO[mydata$SEX=="I"],
     main = 'Histogram of Ratio, Sex = I',
     xlab = 'Ratio',
     col = 'Blue')
hist(mydata$RATIO[mydata$SEX=="M"],
     main = 'Histogram of Ratio, Sex = M',
     xlab = 'Ratio',
     col = 'Red')

# Boxplots
boxplot(mydata$RATIO[mydata$SEX=="F"],
        main = 'Boxplot of Ratio, Sex = F',
        ylab = 'Ratio',
        col = 'Green',
        horizontal = TRUE)
boxplot(mydata$RATIO[mydata$SEX=="I"],
        main = 'Boxplot of Ratio, Sex = I',
        ylab = 'Ratio',
        col = 'Blue',
        horizontal = TRUE)
boxplot(mydata$RATIO[mydata$SEX=="M"],
        main = 'Boxplot of Ratio, Sex = M',
        ylab = 'Ratio',
        col = 'Red',
        horizontal = TRUE)

# QQ Plots and Lines
qqnorm(mydata$RATIO[mydata$SEX=="F"],
       main = 'Normal Q-Q of Ratio, Sex = F',
       col = 'Green')
qqline(mydata$RATIO[mydata$SEX=="F"])
qqnorm(mydata$RATIO[mydata$SEX=="I"],
       main = 'Normal Q-Q of Ratio, Sex = I',
       col = 'Blue')
qqline(mydata$RATIO[mydata$SEX=="I"])
qqnorm(mydata$RATIO[mydata$SEX=="M"],
       main = 'Normal Q-Q of Ratio, Sex = M',
       col = 'Red')
qqline(mydata$RATIO[mydata$SEX=="M"])
```

**Essay Question (2 points): Compare the displays.  How do the distributions compare to normality? Take into account the criteria discussed in the sync sessions to evaluate non-normality.**

***Answer: (The histograms above reveal that for each category of Sex, Ratio has a bell-like distribution that is right-skewed. This is further evidenced by outliers above the whiskers in the boxplots, and the observations above the line at the higher quantiles in the Q-Q plots. Tests of skewness and kurtosis reveal that none of these subsets have skewness = 0 or kurtosis = 3, which in addition to the departure from the line in the Q-Q plots, indicate non-normality. )***

(3)(b) (2 points) The boxplots in (3)(a) indicate that there are outlying RATIOs for each sex. *boxplot.stats()* can be used to identify outlying values of a vector. Present the abalones with these outlying RATIO values along with their associated variables in "mydata". Display the observations by passing a data frame to the kable() function. Basically, we want to output those rows of "mydata" with an outlying RATIO, but we want to determine outliers looking separately at infants, females and males.

```{r Part_3b}
tempsub <- subset(mydata,mydata$RATIO %in% (boxplot.stats(mydata$RATIO)$out))
kable(tempsub)
```

**Essay Question (2 points):  What are your observations regarding the results in (3)(b)?**

***Answer: (The table above indicates several things: 1) Virtually all (except for one) outliers are abalones for which ratio is above the mean. 2) The ages of abalones with outlier values of ratio are concentrated in the classes A1 and A3. Classes A2 and A5 are present as well, but in much smaller numbers. ***

-----

***### Section 4: (8 points) Getting insights about possible predictors.***

(4)(a) (3 points) With "mydata," display side-by-side boxplots for VOLUME and WHOLE, each differentiated by CLASS There should be five boxes for VOLUME and five for WHOLE. Also, display side-by-side scatterplots:  VOLUME and WHOLE versus RINGS. Present these four figures in one graphic:  the boxplots in one row and the scatterplots in a second row. Base R or ggplot2 may be used.


```{r Part_4a}
par(mfrow = c(2,2))
boxplot(mydata$VOLUME ~ mydata$CLASS,
        main = 'Boxplot of Volume by Class',
        ylab = 'Volume',
        xlab = 'Class',
        col = c('Purple','Blue','Green','Yellow','Orange'))

boxplot(mydata$WHOLE ~ mydata$CLASS,
        main = 'Boxplot of Whole by Class',
        ylab = 'Whole',
        xlab = 'Class',
        col = c('Purple','Blue','Green','Yellow','Orange'))

plot(mydata$RINGS,mydata$VOLUME,
     main = 'Scatterplot of Rings vs. Volume',
     xlab = 'Rings',
     ylab = 'Volume')

plot(mydata$RINGS,mydata$WHOLE,
     main = 'Scatterplot of Rings vs. Whole',
     xlab = 'Rings',
     ylab = 'Whole')
```

**Essay Question (5 points) How well do you think these variables would perform as predictors of age?  Explain.**

***Answer: (I do not think that any of these variables would perform well as predictors of age. The boxplots reveal that for both whole weight and volume, the ranges for classes have significant overlap. As a result, whole weight and volume would be both inconsistent and inaccurate predictors of age. This overlap is less present between the younger classes, however, the minimums of all classes' ranges are very similar. The scatterplots reveal the same relationship between rings and whole weight/volume. Abalones with rings equal to or greater than seven in number share virtually the same ranges of whole weight and volume. Similarly, this is less true with abalones with less than seven rings, but the minimum range for volume and whole weight across ring number are similar.)***
-----

***### Section 5: (12 points) Getting insights regarding different groups in the data.***

(5)(a) (2 points) Use *aggregate()* with "mydata" to compute the mean values of VOLUME, SHUCK and RATIO for each combination of SEX and CLASS. Then, using *matrix()*, create matrices of the mean values. Using the "dimnames" argument within *matrix()* or the *rownames()* and *colnames()* functions on the matrices, label the rows by SEX and columns by CLASS. Present the three matrices (Kabacoff Section 5.6.2, p. 110-111). The *kable()* function is useful for this purpose.  You do not need to be concerned with the number of digits presented.

```{r Part_5a}
volumeaggregate <- aggregate(mydata$VOLUME ~ mydata$SEX + mydata$CLASS, FUN  = mean)
volumematrix <- matrix(volumeaggregate[,3],nrow=3,ncol=5)
rownames(volumematrix) <- c('F','I','M')
colnames(volumematrix) <- c('A1','A2','A3','A4','A5')
kable(volumematrix)

shuckaggregate <- aggregate(mydata$SHUCK ~ mydata$SEX + mydata$CLASS, FUN  = mean)
shuckmatrix <- matrix(shuckaggregate[,3],nrow=3,ncol=5)
rownames(shuckmatrix) <- c('F','I','M')
colnames(shuckmatrix) <- c('A1','A2','A3','A4','A5')
kable(shuckmatrix)

ratioaggregate <- aggregate(mydata$RATIO ~ mydata$SEX + mydata$CLASS, FUN  = mean)
ratiomatrix <- matrix(ratioaggregate[,3],nrow=3,ncol=5)
rownames(ratiomatrix) <- c('F','I','M')
colnames(ratiomatrix) <- c('A1','A2','A3','A4','A5')
kable(ratiomatrix)
```

(5)(b) (3 points) Present three graphs. Each graph should include three lines, one for each sex. The first should show mean RATIO versus CLASS; the second, mean VOLUME versus CLASS; the third, mean SHUCK versus CLASS. This may be done with the 'base R' *interaction.plot()* function or with ggplot2 using *grid.arrange()*.

```{r Part_5b, fig.width = 9}
SEX <- mydata$SEX
interaction.plot(mydata$CLASS,SEX,mydata$RATIO,fun = mean,
                 main = 'Mean Value of Ratio for Classes, separated by Sex',
                 ylab = 'Mean Value of Ratio',
                 xlab = 'Class',
                 col = c('Blue','Green','Red'),
                 lty = 1)

interaction.plot(mydata$CLASS,SEX,mydata$VOLUME,fun = mean,
                 main = 'Mean Value of Volume for Classes, separated by Sex',
                 ylab = 'Mean Value of Volume',
                 xlab = 'Class',
                 col = c('Blue','Green','Red'),
                 lty = 1)

interaction.plot(mydata$CLASS,SEX,mydata$SHUCK,fun = mean,
                 main = 'Mean Value of Shuck for Classes, separated by Sex',
                 ylab = 'Mean Value of Shuck',
                 xlab = 'Class',
                 col = c('Blue','Green','Red'),
                 lty = 1)
```

**Essay Question (2 points):  What questions do these plots raise?  Consider aging and sex differences.**

***Answer: (Studying these plots raises the following questions: 1) What accounts for the separation of plot lines between infants and adults? 2) Why does it appear that the plot lines for infants are more linear than those of adults? 3) What is the cause of the change of concavity in the male plot lines for shuck and volume? Why do the females not exhibit the same feature? )***

5(c) (3 points) Present four boxplots using *par(mfrow = c(2, 2)* or *grid.arrange()*. The first line should show VOLUME by RINGS for the infants and, separately, for the adult; factor levels "M" and "F," combined. The second line should show WHOLE by RINGS for the infants and, separately, for the adults. Since the data are sparse beyond 15 rings, limit the displays to less than 16 rings. One way to accomplish this is to generate a new data set using subset() to select RINGS < 16.  Use ylim = c(0, 1100) for VOLUME and ylim = c(0, 400) for WHOLE.  If you wish to reorder the displays for presentation purposes or use ggplot2 go ahead.

```{r Part_5c}
par(mfrow = c(2, 2))
boxplot(mydata$VOLUME[mydata$SEX == 'I'] ~ mydata$RINGS[mydata$SEX == 'I'],
        subset = mydata$RINGS <= 16,
        main = 'Boxplot of Volume by Rings, Sex = "I"',
        xlab = 'Rings',
        ylab = 'Volume')

boxplot(mydata$VOLUME[mydata$SEX != "I"] ~ mydata$RINGS[mydata$SEX != "I"],
        subset = mydata$RINGS <= 16,
        main = 'Boxplot of Volume by Rings, Sex = "M/F"',
        xlab = 'Rings',
        ylab = 'Volume')

boxplot(mydata$WHOLE[mydata$SEX == 'I'] ~ mydata$RINGS[mydata$SEX == 'I'],
        subset = mydata$RINGS <= 16,
        main = 'Boxplot of Whole by Rings, Sex = "I"',
        xlab = 'Rings',
        ylab = 'Whole')

boxplot(mydata$WHOLE[mydata$SEX != "I"] ~ mydata$RINGS[mydata$SEX != "I"],
        subset = mydata$RINGS <= 16,
        main = 'Boxplot of Whole by Rings, Sex = "M/F"',
        xlab = 'Rings',
        ylab = 'Whole')

```

**Essay Question (2 points):  What do these displays suggest about abalone growth?  Also, compare the infant and adult displays.  What differences stand out?**

***Answer: (These displays suggest that as rings (age) increase, abalones generally grow in size (as measured by whole weight and volume.) However, infant abalones demonstrate a pattern of means more akin to a parabola, in which values of whole weight and volume increase until rings equal about 12, after which they tend to fall. These displays also demonstrate consistent patterns of variation in these two variables as rings (age) increase. Variations of these two variables among abalones with fewer rings is signicantly smaller than it is for abalones with ring counts of seven and above. However, this variation appears to shrink among abalones with the greatest values of rings (age.))***

-----

***### Section 6: (11 points) Conclusions from the Exploratory Data Analysis (EDA).***


**Conclusions**


**Essay Question 1) (5 points)	Based solely on these data, what are plausible statistical reasons that explain the failure of the original study? Consider to what extent physical measurements may be used for age prediction.**

***Answer: (I believe the main reason that this study failed to help develop an accurate and reliable method of estimating abalone age is that the variables they collected were simply not good predictors of age. As demonstrated above, outside of the age span of zero to (roughly) seven, there is consistent and large overlap among classes for all variables used to measure physical characteristics of abalones. This means that estimations of age based on these variables will inaccurate and inconsistent. The resulting conclusion is that the wrong variables were measured. Additionally, the distributions of the variables they collected had outliers, skewness, and kurtosis in values that indicate non-normality. While we have not used descriptive statistics of these variables for the purpose of inferential statistics, these features have negative statistical implications.)***

**Essay Question 2) (3 points) Do not refer to the abalone data or study.  If you were presented with an overall histogram and summary statistics from a sample of some population or phenomenon and no other information, what questions might you ask before accepting them as representative of the sampled population or phenomenon?**

***Answer: (The first pieces of information I would ask for would pertain to the sampling frame (what it was defined as, what its size was, what the characteristics of interest were, how they were distributed). Additionally, if there were any summary statistics available on the population, I would like have them. I would then ask about how observations and variables were defined (as well as how variables were measured.) I would also like to know the size and composition of the sample and what method was used to draw it. If I came to the conclusions that the sampling frame was well-defined, the sample was drawn appropriately and with sufficient size, the histogram and summary statistics were not immediately indicative of non-normality, and comparison of the available statistics seemed reasonable, I would be willing to accept them as representative... after completing legitimate statistical tests for normality on the sample.)***

**Essay Question 3)  (3 points)	Do not refer to the abalone data or study.  What do you see as difficulties analyzing data derived from observational studies? Can causality be determined?  What might be learned from such studies?**

***Answer: (The largest difficulty I imagine one would encounter while using observational data from a study would be the validity of the data. I mean this first in the statistical sense (i.e. how well the sampling frame, observations, and variables are defined, how representative of the frame the sample is, potentially the assumptions that can or cannot be made given sample size and variable values.) I also mean this in a more literal way, as observational data can be prone to error and the methods through which variables are measured may not be appropriate or accurate. Another difficulty I would image arises is the issue of using data that is not designed specifically for your purpose, as the data only has the potential to be the most relevant and useful data if you are using it for the same exact purpose as the study. I would imagine that the data requirements for the mathematical and statistical methods used to determine causality would be far beyond the scope of data from observational studies. Like basic analyses for descriptive and inferential statistics, they could certainly be tried, but I would expect the probability of validity to be quite small. Depending on the data, I imagine one or more of several insights could be learned. These would include whether or not you agree with the results and conclusions of the study, features of the sample based on descriptive statistics, estimates of population features given your own analyses of the data, as well as any insights that could come from combining the data with other sources for whatever purpose one has.)***
